---
title: "Suchan's_lovely_R_notebook♥"
output: html_notebook
---

# **2.7 Coercion**

In general, *coercion* is an attempt by R to be flexible with data types. When an entry does not match the expected, some of the prebuilt R functions try to guess what was meant before throwing an error. This can also lead to confusion. Failing to understand *coercion* can drive programmers crazy when attempting to code in R since it behaves quite differently from most other languages in this regard. Let’s learn about it with some examples.

We said that vectors must be all of the same type. So if we try to combine, say, numbers and characters, you might expect an error:

```{r}
x <- c(1, "canada", 3)
```

But we don’t get one, not even a warning! What happened? Look at ```x``` and its class:
```{r}
x

class(x)
```

R *coerced* the data into characters. It guessed that because you put a character string in the vector, you meant the 1 and 3 to actually be character strings ```"1"``` and ```“3”```. The fact that not even a warning is issued is an example of how coercion can cause many unnoticed errors in R.

R also offers functions to change from one type to another. For example, you can turn numbers into characters with:
```{r}
x <- 1:5
y <- as.character(x)
y
```

You can turn it back with ```as.numeric```:
```{r}
as.numeric(y)
```

This function is actually quite useful since datasets that include numbers as character strings are common.


## **2.7.1 Not availables (NA)**

When a function tries to coerce one type to another and encounters an impossible case, it usually gives us a warning and turns the entry into a special value called an ```NA``` for “not available”. For example:
```{r}
x <- c("1", "b", "3")
as.numeric(x)
```

R does not have any guesses for what number you want when you type ```b```, so it does not try.

As a data scientist you will encounter the ```NA```s often as they are generally used for missing data, a common problem in real-world datasets.


# **2.8 Exercises**

1. Use the function ```c``` to create a vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto, which are 35, 88, 42, 84, 81, and 30 degrees Fahrenheit. Call the object ```temp```.
```{r}
temp <- c(35,88,42,84,81,30)
```

2. Now create a vector with the city names and call the object ```city```.
```{r}
city <- c("Beijing","Lagos","Paris","Rio de Janerio","San Juan","Toronto")
```

3. Use the ```names``` function and the objects defined in the previous exercises to associate the temperature data with its corresponding city.
```{r}
names(temp) <- city
temp
```

4. Use the ```[``` and ```:``` operators to access the temperature of the first three cities on the list.
```{r}
temp[1:3]
```

5. Use the ```[``` operator to access the temperature of Paris and San Juan.
```{r}
temp[c(3,5)]
```

6. Use the ```:``` operator to create a sequence of numbers  
$12, 13, 14,..., 73$.
```{r}
12:73
```

7. Create a vector containing all the positive odd numbers smaller than 100.
```{r}
c(seq(1, 100, 2))
```

8. Create a vector of numbers that starts at 6, does not pass 55, and adds numbers in increments of 4/7: 6, 6 + 4/7, 6 + 8/7, and so on. How many numbers does the list have? Hint: use ```seq``` and ```length```
```{r}
V <- c(seq(6,55,4/7))
length(V)
```

9. What is the class of the following object ```a <- seq(1, 10, 0.5)```?
```{r}
a <- seq(1, 10, 0.5)
class(a)
```

10. What is the class of the following object ```a <- seq(1, 10)```?
```{r}
a <- seq(1,10)
class(a)
```

11. The class of ```class(a<-1)``` is numeric, not integer. R defaults to numeric and to force an integer, you need to add the letter ```L```. Confirm that the class of ```1L``` is integer.
```{r}
class(a <- 1L)
```

12. Define the following vector:
```{r}
x <- c("1", "3", "5")
```
and coerce it to get integers.
```{r}
X <- as.integer(x)
class(X)

```

# **2.9 Sorting**

Now that we have mastered some basic R knowledge, let’s try to gain some insights into the safety of different states in the context of gun murders.

## **2.9.1 ```sort```**

Say we want to rank the states from least to most gun murders. The function ```sort``` sorts a vector in increasing order. We can therefore see the largest number of gun murders by typing:
```{r}
library(dslabs)
data(murders)
sort(murders$total)
```

However, this does not give us information about which states have which murder totals. For example, we don’t know which state had 1257.

## **2.9.2 ```order```**

The function ```order``` is closer to what we want. It takes a vector as input and returns the vector of indexes that sorts the input vector. This may sound confusing so let’s look at a simple example. We can create a vector and sort it:
```{r}
x <- c(31,4,15,92,65)
sort(x)
```

Rather than sort the input vector, the function ```order``` returns the index that sorts input vector:
```{r}
index <- order(x)
x[index]
```

This is the same output as that returned by ```sort(x)```. If we look at this index, we see why it works:
```{r}
x
order(x)
```

The second entry of ```x``` is the smallest, so ```order(x)``` starts with ```2```. The next smallest is the third entry, so the second entry is ```3``` and so on.

How does this help us order the states by murders? First, remember that the entries of vectors you access with ```$``` follow the same order as the rows in the table. For example, these two vectors containing state names and abbreviations, respectively, are matched by their order:
```{r}
murders$state[1:6]
murders$abb[1:6]
```

This means we can order the state names by their total murders. We first obtain the index that orders the vectors according to murder totals and then index the state names vector:
```{r}
ind <- order(murders$total)
murders$abb[ind]
```

According to the above, California had the most murders.

## **2.9.3 ```max``` and ```which.max```**

If we are only interested in the entry with the largest value, we can use ```max``` for the value:
```{r}
max(murders$total)

```

and ```which.max``` for the index of the largest value:
```{r}
i_max <- which.max(murders$total)
murders$state[i_max]
```

For the minimum, we can use ```min``` and ```which.min``` in the same way.

Does this mean California is the most dangerous state? In an upcoming section, we argue that we should be considering rates instead of totals. Before doing that, we introduce one last order-related function: ```rank```.

## **2.9.4 ```rank```**

Although not as frequently used as ```order``` and ```sort```, the function ```rank``` is also related to order and can be useful. For any given vector it returns a vector with the rank of the first entry, second entry, etc., of the input vector. Here is a simple example:
```{r}
x <- c(31,4,15,92,65)
rank(x)
```

To summarize, let’s look at the results of the three functions we have introduced:
```{r}
a <- x
b <- sort(x)
c <- order(x)
d <- rank(x)
content <- c(a,b,c,d)
mat <- matrix(content, nrow=5,ncol=4)
o <- c("original","sort","order","rank")
O <- data.frame(mat)
colnames(O) <- o
O

```

## **2.9.5 Beware of recycling**

Another common source of unnoticed errors in R is the use of *recycling*. We saw that vectors are added elementwise. So if the vectors don’t match in length, it is natural to assume that we should get an error. But we don’t. Notice what happens:
```{r}

```

























