---
title: "Suchan's_R_notebook♥"
output: html_notebook
---

# **Chapter4 The tidyverse**

Up to now we have been manipulating vectors by reordering and subsetting them through indexing. However, once we start more advanced analyses, the preferred unit for data storage is not the vector but the data frame. In this chapter we learn to work directly with data frames, which greatly facilitate the organization of information. We will be using data frames for the majority of this book. We will focus on a specific data format referred to as *tidy* and on specific collection of packages that are particularly helpful for working with *tidy* data referred to as the *tidyverse*.


We can load all the tidyverse packages at once by installing and loading the **tidyverse** package:
```{r}
library(tidyverse)
```

We will learn how to implement the tidyverse approach throughout the book, but before delving into the details, in this chapter we introduce some of the most widely used tidyverse functionality, starting with the **dplyr** package for manipulating data frames and the **purrr** package for working with functions. Note that the tidyverse also includes a graphing package, **ggplot2**, which we introduce later in Chapter 7 in the Data Visualization part of the book; the **readr** package discussed in Chapter 5; and many others. In this chapter, we first introduce the concept of *tidy data* and then demonstrate how we use the tidyverse to work with data frames in this format.


## **4.1 Tidy data**

We say that a data table is in tidy format if each row represents one observation and columns represent the different variables available for each of these observations. The ```murders``` dataset is an example of a tidy data frame.
```{r}
library(dslabs)
head(murders)
```

Each row represent a state with each of the five columns providing a different variable related to these states: name, abbreviation, region, population, and total murders.

To see how the same information can be provided in different formats, consider the following example:
```{r}
#>       country year fertility
#> 1     Germany 1960      2.41
#> 2 South Korea 1960      6.16
#> 3     Germany 1961      2.44
#> 4 South Korea 1961      5.99
#> 5     Germany 1962      2.47
#> 6 South Korea 1962      5.79

```

This tidy dataset provides fertility rates for two countries across the years. This is a tidy dataset because each row presents one observation with the three variables being country, year, and fertility rate. However, this dataset originally came in another format and was reshaped for the **dslabs** package. Originally, the data was in the following format:
```{r}
#>       country 1960 1961 1962
#> 1     Germany 2.41 2.44 2.47
#> 2 South Korea 6.16 5.99 5.79

```


The same information is provided, but there are two important differences in the format: 1) each row includes several observations and 2) one of the variables, year, is stored in the header. For the tidyverse packages to be optimally used, data need to be reshaped into ```tidy``` format, which you will learn to do in the Data Wrangling part of the book. Until then, we will use example datasets that are already in tidy format.


Although not immediately obvious, as you go through the book you will start to appreciate the advantages of working in a framework in which functions use tidy formats for both inputs and outputs. You will see how this permits the data analyst to focus on more important aspects of the analysis rather than the format of the data.

## **4.2 Exercises**

1. Examine the built-in dataset ```co2```. Which of the following is true:

a. ```co2``` is tidy data: it has one year for each row.
b. ```co2``` is not tidy: we need at least one column with a character vector.
c. ```co2``` is not tidy: it is a matrix instead of a data frame.
d. ```co2``` is not tidy: to be tidy we would have to wrangle it to have three columns(year, month and value), then each co2 observation would have a row.
```{r}
co2

#>answer is 'd'
```

2. Examine the built-in dataset ```ChickWeight```. Which of the following is true:

a. ```ChickWeight``` is not tidy: each chick has more than one row.
b. ```ChickWeight``` is tidy: each observation(a weight) is represented by one row. The chick from which this measurement came is one of the varialbes.
c. ```ChickWeight``` is not tidy: we are missing the year column.
d. ```ChickWeight``` is tidy: it is stored in a data frame.
```{r}
ChickWeight

#>answer is 'd'
```

3. Examine the built-in dataset ```BOD```. Which of the following is true:

a. ```BOD``` is not tidy: it only has six rows.
b. ```BOD``` is not tidy: the first column is just an index.
c. ```BOD``` is tidy: each row is an observation with two values(time and demand)
d. ```BOD``` is tidy: all small datasets are tidy by definition.


```{r}
BOD

#>the answer is 'c'

```

4. Which of the following built-in datasets is tidy (you can pick more than one):

a. ```BJsales```
b. ```EuStockMarkets```
c. ```DNase```
d. ```Formaldehyde```
e. ```Orange```
f. ```UCBAdmissions```
```{r}
BJsales
EuStockMarkets
DNase
Formaldehyde
Orange
UCBAdmissions

#>the answer is 'b','c','d','e'
```


## **4.3 Manipulating data frames**

The **dplyr** package from the **tidyverse** introduces functions that perform some of the most common operations when working with data frames and uses names for these functions that are relatively easy to remember. For instance, to change the data table by adding a new column, we use ```mutate```. To filter the data table to a subset of rows, we use ```filter```. Finally, to subset the data by selecting specific columns, we use ```select```.

### **4.3.1 Adding a column with ```mutate```**

We want all the necessary information for our analysis to be included in the data table. So the first task is to add the murder rates to our murders data frame. The function ```mutate``` takes the data frame as a first argument and the name and values of the variable as a second argument using the convention ```name = values```. So, to add murder rates, we use:
```{r}
library(dslabs)
data('murders')
murders <- mutate(murders, rate=total/population*100000)
```

Notice that here we used ```total``` and ```population``` inside the function, which are objects that are **not** defined in our workspace. But why don’t we get an error?

This is one of **dplyr**’s main features. Functions in this package, such as ```mutate```, know to look for variables in the data frame provided in the first argument. In the call to mutate above, ```total``` will have the values in ```murders$total```. This approach makes the code much more readable.

We can see that the new column is added:
```{r}
head(murders)
```

Although we have overwritten the original ```murders``` object, this does not change the object that loaded with ```data(murders)```. If we load the ```murders``` data again, the original will overwrite our mutated version.


### **Subsetting with ```filter```**

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the ```filter``` function, which takes the data table as the first argument and then the conditional statement as the second. Like ```mutate```, we can use the unquoted variable names from ```murders``` inside the function and it will know we mean the columns and not objects in the workspace.
```{r}
filter(murders, rate<=0.71)
```

### **4.3.3 Selecting columns with ```select```**

Although our data table only has six columns, some data tables include hundreds. If we want to view just a few, we can use the **dplyr** ```select``` function. In the code below we select three columns, assign this to a new object and then filter the new object:
```{r}
new_table <- select(murders,state,region,rate)
filter(new_table,rate<=0.71)
```


In the call to ```select```, the first argument ```murders``` is an object, but ```state```, ```region```, and ```rate``` are variable names.


## **4.4 Exercises**

























